Edge sign prediction in general graphs and Correlation Clustering

### 2017-11-28 cc/cc_variants.tex
add some ref
add 2 refs
rewrite related troll works
add some refs
tiny fixes and ref all over the places
typos and stability in cc

% leftover:
% - something about fuzzy edges, ie we're not sure they are in the graph or not (not super
%   interesting in my opinion) Clustering with partial information, Theoretical Computer Science,
%   Volume 411, Issue 7, 2010, Pages 1202-1211, ISSN 0304-3975,
%   \url{http://dx.doi.org/10.1016/j.tcs.2009.12.016}
% - \paragraph{Hypergraphs} \Textcite{Kim2011} show a LP relaxation on hypergraph and
%   \textcite{Ricatte13} describe a class of hypergraphs that can be reduced to signed graph.  Jörg
%   Hendrik Kappes, Markus Speth, Gerhard Reinelt, Christoph Schnörr, Higher-order segmentation via
%   multicuts, Computer Vision and Image Understanding, Volume 143, 2016, Pages 104-119, ISSN
%   1077-3142, \url{http://dx.doi.org/10.1016/j.cviu.2015.11.005}
% - Interactive \pcc{}?~\autocite{interactiveCC16}, could it go into active?  Geerts, F. & Ndindi,
%   R. Bounded correlation clustering. Int. J. Data Sci. Anal. 1, 17–35 (2016)
%   \url{https://doi.org/10.1007/s41060-016-0005-2}

So far we focused on \pcc{} in its rawest form, that is solving the \mind{} and \maxa{} objectives
in the case where the general binary-labeled graph is known in advance. We will now see first some
special cases, namely when the weights obey the triangle inequality (to solve \msc{}) or when the
graph is bipartite and then move to variants. We will consider more general objectives, when the
edges are labeled categorically instead of binary, when nodes can belong simultaneously to several
clusters or when we optimize local objectives per nodes instead of global ones. Finally we will also look at
clustering in signed graphs in general, using spectral methods or heuristics from the community
detection literature.

@@

In \msc{}, the goal is to output a clustering which best summarizes (or agrees with) the several
given input clusterings of the same set of objects.  Motivations include robustness ---by using an
ensemble of clusterings from diverse methods--- and privacy ---if the clusterings were computed by
different parties each considering only a subset of the objects attributes. We can build the
complete graph of these objects, with weights set to the fraction of clusterings that place two
objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
first show by \textcite{EarlyConsensusClustering03} , finding the optimal clustering is therefore
an instance of \pcc{} where the weights obey the triangular inequality. \textcite{Gionis2007} give a
deterministic $3$-approximation using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
the minimization version is \APXh{}, even when the input is made of three clusterings and give a
combinatorial $\nicefrac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
conducted by \textcite{Bertolacci07} and \textcite{Filkov08}. The former describe a scalable
approach that first samples a small portion of the data, runs a (potentially computationally
expensive) approximation algorithm and finally augment the resulting partition by adding to it the
unsampled nodes one by one.  Experiments confirm that the running time is greatly improved compared
with the linear program methods while the resulting objective value is essentially the same. Note,
however, that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
If we have $k$ input clusterings $\cluster^1,\ldots,\cluster^k$ and we parameterized the problem by
$t$, which is the sum over the input clusterings of the number of pairs of objects that are clustered
differently by a solution $\cluster^\star$ and $\cluster^i$, then there is a polynomial algorithm
running in $O(4.24^{\nicefrac{t}{k}}\cdot \nicefrac{t}{k}^3 +
kn^2)$~\autocite{parameterizedConsensus14}.

@@

In the so called \textsc{Chromatic}-\pcc{} setting, \enquote{positive} edges are now associated with
one of $L$ possible colors and the goal is to form clusters mostly made up of edges with the one same
color. Namely, a disagreement is now a negative edge between clusters or a within-cluster edge
whose color differs from the majority color of that cluster. This is motivated by edge-labeled graph
in social networks, biology and citation networks and will discuss such applications in
\autoref{chap:vector}. As a generalisation of \pcc{}, it is \NPc{} but \textcite{Bonchi2012a}
present a modification of the \ccpivot{} algorithm that pick edges instead of nodes as pivots, and
grow clusters by adding monochromatic triangles. This gives an approximation factor of six times the
maximum degree of the graph. They also present a method when the number $k$ of clusters is fixed
beforehand, starting with an initial partition and then alternating between finding the majority
color of clusters and finding better clusters. Unfortunately, the maximum degree of a graph can be
as large as $n$. However, \textcite{Anava2015} present constant factor approximations. Namely, they
show the problem can be reduced to classical \pcc{} by setting all edges incident to a node $u$ to
negative if they are not of the majority color of $u$. They then apply the regular \ccpivot{} and
show this gives an $11$-approximation to the original problem. Furthermore, they also write a linear
program and round it using the \regionGrow{} method to obtain an
approximation factor of $4$. \Textcite{multiChromatic15} extend this line of work to the case were a
single edge can carry a \emph{set} of labels and adapt their randomized algorithm so that the
approximation factor is multiplied by the size of the input label set.

@@

\Textcite{pmlr-v48-puleo16} also modify the \mind{} objective to make it more general. Based on the
classic \pcc{} linear program, they define a \enquote{\emph{fractional clustering} of $G$ as a
vector $x$ indexed by $V$ such that $x_{uv} \in [0, 1]$ for all $uv \in \binom{V}{2}$ and such that
$x_{vz} \leq x_{vw} + x_{wz}$ for all distinct $v, w, z \in V$}. They also define \enquote{The
error vector $err(x)$ of $x$, as a real vector indexed by $V$ whose coordinates are}
\begin{equation*}
  \mathrm{err}(x)_u = \sum_{v\in\nei^+(u)} x_{uv} + \sum_{v\in\nei^-(u)} (1-x_{uv})
\end{equation*}

@@

\paragraph{Spectral Clustering}

@@

Finally, \textcite{mGeometricMean16} show that the Laplacians defined so far can be seen as
arithmetic means of the Laplacian $L^+$ of the positive subgraph $G^+=(V, E^+)$ and the signless
Laplacian $Q^-$ of the negative subgraph $G^-=(V, E^-)$, where $Q^- = D^- + A^-$. They suggest
instead to use a geometric mean defined for two positive matrices $A$ and $B$ as $A\#B =
A^{\shalf{}} \left(A^{-\shalf{}} BA^{-\shalf{}} \right)^{\shalf{}} A^{\shalf{}}$. This suggestion is
based on the fact that if $u$ is a common eigenvector of both $A$ and $B$ with eigenvalue $\lambda$
and $\mu$ respectively, then $u$ is an eigenvector of $A+B$ with eigenvalue $\lambda+\mu$ and an
eigenvector of $A\#B$ with eigenvalue $\sqrt{\lambda\mu}$. Therefore, the $k$ smallest eigenvalues
of the geometric mean Laplacian will be influenced by both smallest eigenvalues of $L^+$
(corresponding to assortative clusters in $G^+$) and of $Q^-$ (corresponding to disassortative
clusters in $G^-$), while this is note the case for the arithmetic mean of Laplacians.

@@

The clustering problem is often named community detection in the context of social networks, and
several methods developed by practitioners have been extended to signed graphs. While they do not
necessarily considered the \pcc{} objective, and especially not its optimum, we still give a brief
overview of them, as they tend to have been more tested experimentally. For instance, to find the
cluster of node $u$, \textcite{Yang2007} use a random walk approach on the positive subgraph $G^+$
to compute the probability of each node to reach $u$ in $T$ steps, sort the nodes accordingly and
then find a threshold  based on the number of disagreements.
The one node moves local heuristic that we described in the Physics-inspired paragraph
\vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
minimize the number of disagreements and maximize a signed variant of the
modularity~\autocites{Li2013}{Amelio2013}. \Textcite{Anchuri2012} also consider these two objectives
by seeing them as eigenvalue problems and devise a iterative splitting procedure.
The overlapping community detection variant is considered by \textcite{Chen14}, who used a signed
probabilistic mixture model. Namely, an edge selects a pair of cluster $r,s$ with probability
$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and chooses its endpoints
$u$ and $v$ with probability $\theta_{ru}$ and $\theta_{sv}$. $\theta_{ru}$ is therefore the soft
membership of node $u$ to cluster $r$, and those parameters are estimated using the
expectation-maximization algorithm. The same model is extended to directed graphs by
\textcite{Jiang2015}, who strangely enough names it stochastic blockmodel, although the focus is
still on edge and not nodes.
In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
positive edges and densely connected by negative edges to each other. Thee dub such subgraph
\emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
problem as a constrained quadratic optimization, they propose a faster iterative local search
heuristic.
% relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}

>>>>

% leftover:
% - something about fuzzy edges, ie we're not sure they are in the graph or not (not super
%   interesting in my opinion) Clustering with partial information, Theoretical Computer Science,
%   Volume 411, Issue 7, 2010, Pages 1202-1211, ISSN 0304-3975,
%   \url{http://dx.doi.org/10.1016/j.tcs.2009.12.016}
% - \paragraph{Hypergraphs} \Textcite{Kim2011} show a LP relaxation on hypergraph and
%   \textcite{Ricatte13} describe a class of hypergraphs that can be reduced to signed graph.  Jörg
%   Hendrik Kappes, Markus Speth, Gerhard Reinelt, Christoph Schnörr, Higher-order segmentation via
%   multicuts, Computer Vision and Image Understanding, Volume 143, 2016, Pages 104-119, ISSN
%   1077-3142, \url{http://dx.doi.org/10.1016/j.cviu.2015.11.005}
% - Interactive \pcc{}?~\autocite{interactiveCC16}, could it go into active?  Geerts, F. & Ndindi,
%   R. Bounded correlation clustering. Int. J. Data Sci. Anal. 1, 17–35 (2016)
%   \url{https://doi.org/10.1007/s41060-016-0005-2}
% - CC with triplet constraint instead of signs \autocite{relativeCC17}

So far we focused on \pcc{} in its rawest form, that is solving the \mind{} and \maxa{} objectives
in the case where the general binary-labeled graph is known in advance. We will now see first some
special cases, namely when the weights obey the triangle inequality (to solve \msc{}) or when the
graph is bipartite and then move to variants. We will consider more general objectives, when the
edge labels are categorical instead of binary, when nodes can belong simultaneously to several
clusters or when we optimize local objectives per nodes instead of global ones. Finally we will also look at
clustering in signed graphs in general, using spectral methods or heuristics from the community
detection literature.

@@

In \msc{}, the goal is to output a clustering which best summarizes (or agrees with) the several
given input clusterings of the same set of objects.  Motivations include robustness ---by using an
ensemble of clusterings from diverse methods--- and privacy ---if the clusterings were computed by
different parties each considering only a subset of the objects attributes. We can build the
complete graph of these objects, with weights set to the fraction of clusterings that place two
objects in different clusters, thus representing a kind of distance in the space of clusterings.  As
first show by \textcite{EarlyConsensusClustering03} , finding the optimal clustering is therefore
an instance of \pcc{} where the weights obey the triangular inequality. \textcite{Gionis2007} give a
deterministic $3$-approximation using the \regionGrow{} method. Later \textcite{Bonizzoni2008} show that
the minimization version is \APXh{}, even when the input is made of only three clusterings and give a
combinatorial $\nicefrac{4}{5}$-approximation for the maximization problem. Experimental evaluations are
conducted by \textcite{Bertolacci07} and \textcite{Filkov08}. The former describe a scalable
approach that first samples a small portion of the data, runs a (potentially computationally
expensive) approximation algorithm and finally augment the resulting partition by adding to it the
unsampled nodes one by one.  Experiments confirm that the running time is greatly improved compared
with the linear program methods while the resulting objective value is essentially the same. Note,
however, that LP methods can be applied in practice thanks to some tricks~\autocite{ConsensusLP10}.
If we have $k$ input clusterings $\cluster^1,\ldots,\cluster^k$ and we parameterized the problem by
$t$, which is the sum over the input clusterings of the number of pairs of objects that are clustered
differently by a solution $\cluster^\star$ and $\cluster^i$, then there is a polynomial algorithm
running in $O(4.24^{\nicefrac{t}{k}}\cdot \nicefrac{t}{k}^3 +
kn^2)$~\autocite{parameterizedConsensus14}.

@@

In the so called \textsc{Chromatic}-\pcc{} setting, \enquote{positive} edges are now associated with
one of $L$ possible colors and the goal is to form clusters mostly made up of edges with the one same
color. Namely, a disagreement is now a negative edge between clusters or a within-cluster edge
whose color differs from the majority color of that cluster. This is motivated by edge-labeled graph
in social networks, biology and citation networks and will discuss such applications in
\autoref{chap:vector}. As a generalisation of \pcc{}, it is \NPc{} but \textcite{Bonchi2012a}
present a modification of the \ccpivot{} algorithm that pick edges instead of nodes as pivots, and
grow clusters by adding monochromatic triangles. This gives an approximation factor of six times the
maximum degree of the graph. They also present a method when the number $k$ of clusters is fixed
beforehand, starting with an initial partition and then alternating between finding the majority
color of clusters and finding better clusters. An improved heuristic algorithm is given
in~\autocite{CCCheuristic16}. Unfortunately, the maximum degree of a graph can be
as large as $n$. However, \textcite{Anava2015} present constant factor approximations. Namely, they
show the problem can be reduced to classical \pcc{} by setting all edges incident to a node $u$ to
negative if they are not of the majority color of $u$. They then apply the regular \ccpivot{} and
show this gives an $11$-approximation to the original problem. Furthermore, they also write a linear
program and round it using the \regionGrow{} method to obtain an
approximation factor of $4$. \Textcite{multiChromatic15} extend this line of work to the case were a
single edge can carry a \emph{set} of labels and adapt their randomized algorithm so that the
approximation factor is multiplied by the size of the input label set.

@@

\Textcite{pmlr-v48-puleo16} also modify the \mind{} objective to make it more general. Based on the
classic \pcc{} linear program, they define a \enquote{\emph{fractional clustering} of $G$ as a
vector $x$ indexed by $V$ such that $x_{uv} \in [0, 1]$ for all $uv \in \binom{V}{2}$ and such that
$x_{vz} \leq x_{vw} + x_{wz}$ for all distinct $v, w, z \in V$}. They also define \enquote{the
error vector $err(x)$ of $x$, as a real vector indexed by $V$ whose coordinates are}
\begin{equation*}
  \mathrm{err}(x)_u = \sum_{v\in\nei^+(u)} x_{uv} + \sum_{v\in\nei^-(u)} (1-x_{uv})
\end{equation*}

@@

\paragraph{Spectral Clustering}
\label{par:cc_spectral}

@@

Finally, \textcite{mGeometricMean16} show that the Laplacians defined so far can be seen as
arithmetic means of the Laplacian $L^+$ of the positive subgraph $G^+=(V, E^+)$ and the signless
Laplacian $Q^-$ of the negative subgraph $G^-=(V, E^-)$, where $Q^- = D^- + A^-$. They suggest
instead to use a geometric mean defined for two positive matrices $A$ and $B$ as $A\#B =
A^{\shalf{}} \left(A^{-\shalf{}} BA^{-\shalf{}} \right)^{\shalf{}} A^{\shalf{}}$. This suggestion is
based on the fact that if $u$ is a common eigenvector of both $A$ and $B$ with eigenvalue $\lambda$
and $\mu$ respectively, then $u$ is an eigenvector of $A+B$ with eigenvalue $\lambda+\mu$ and an
eigenvector of $A\#B$ with eigenvalue $\sqrt{\lambda\mu}$. Therefore, the $k$ smallest eigenvalues
of the geometric mean Laplacian will be influenced by both smallest eigenvalues of $L^+$
(corresponding to assortative clusters in $G^+$) and of $Q^-$ (corresponding to disassortative
clusters in $G^-$), while this is not the case for the arithmetic mean of Laplacians.

@@

The clustering problem is often named community detection in the context of social networks, and
several methods developed by practitioners have been extended to signed graphs. While they do not
necessarily considered the \pcc{} objective, and especially not its optimum, we still give a brief
overview of them, as they tend to have been more tested experimentally. For instance, to find the
cluster of node $u$, \textcite{Yang2007} use a random walk approach on the positive subgraph $G^+$
to compute the probability of each node to reach $u$ in $T$ steps, sort the nodes accordingly and
then find a threshold  based on the number of disagreements.
The one node moves local heuristic that we described in the Physics-inspired paragraph
\vpageref{par:cc_physics} can also be formulated as genetic algorithms that simultaneously try to
minimize the number of disagreements and maximize a signed variant of the
modularity~\autocites{Li2013}{Amelio2013}. \Textcite{Anchuri2012} also consider these two objectives
by seeing them as eigenvalue problems and devise a iterative splitting procedure.
The overlapping community detection variant is considered by \textcite{Chen14}, who used a signed
probabilistic mixture model. Namely, an edge selects a pair of cluster $r,s$ with probability
$\omega_{rs}$ (where $r=s$ if the edge is positive and $r\neq s$ otherwise) and chooses its endpoints
$u$ and $v$ with probability $\theta_{ru}$ and $\theta_{sv}$. $\theta_{ru}$ is therefore the soft
membership of node $u$ to cluster $r$, and those parameters are estimated using the
expectation-maximization algorithm. The same model is extended to directed graphs by
\textcite{Jiang2015}, who strangely enough names it stochastic blockmodel, although the focus is
still on edge and not nodes.
In a similar spirit to \maxa{}$[k]$, \textcite{SignedGang} focus on finding $k$ subgraphs dense in
positive edges and densely connected by negative edges to each other. They dub such subgraph
\emph{Oppositive Cohesive Groups}, or more vividly \emph{Gangs in War}, and after formulating the
problem as a constrained quadratic optimization, they propose a faster iterative local search
heuristic. When $k=2$, these subgraphs are called antagonistic communities and a specific data
mining approach was proposed by \textcite{quasiGang16}.
% relaxed structural balance \autocite{Doreian2009} but used the same approach as \textcite{Early96}

### 2017-11-28 cc/gtx_related.tex
typos and stability in cc

\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
for instance the Elkin paper \enquote{Our centralized randomized algorithm computes (with
probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
n}{n}\right)\right)$ edges in $O(|E|)$ time.

>>>>

\url{http://www.siam.org/meetings/da17/schedule.html} SODA 13B \url{http://dl.acm.org/citation.cfm?id=3039686}
for instance the Elkin paper~\autocite{Spanner17} \enquote{Our centralized randomized algorithm computes (with
probability close to 1), a $(2k - 1)$-spanner with $n \cdot (1 + O(\frac{\log k}{n}))$ edges in
$O(|E|)$ time, whenever $k = \Omega(\log n)$. Note that when $k = \omega(\log n)$, the number of
edges is $n(1+o(1))$, i.e., in this range the algorithm computes an ultra-sparse spanner in $O(|E|)$
time.} For instance, if $k=5\log n$, we get a $10\log n$-spanner with $n\left(1+O\left(\frac{\log\log
n}{n}\right)\right)$ edges in $O(|E|)$ time.
